Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon

### 2022-01-29 paper.Rmd
finished working

---
title: "Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon"
output:
  bookdown::word_document2:
    reference_docx: wordstyles.docx
    toc: true
    toc_float: true
    number_sections: true
    css: custom.css
  #  code_folding: hide
always_allow_html: true
bibliography: sources.bib

@@

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(report)
library(lmerTest)
library(plotly)
library(corrr)
library(boot)
library(zeallot)
library(pander)
library(sjPlot)
library(ggeffects)
library(gridExtra)
library(gghalves)
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message = FALSE)
```

@@

Iconicity--defined here as a perceived resemblance between aspects of form and meaning--is now increasingly recognised as a *key property of language* [see e.g. @perniss_iconicity_2010; @perniss_bridge_2014; @dingemanse_arbitrariness_2015 for a review], raising questions about where iconic mappings come from, and what they do. To answer these questions, we first need to be able to *identify* iconicity in language and, ideally, to *measure* it.

@@

With descriptive approaches, iconicity as a theoretical concept is coded for directly (e.g. @pietrandrea_iconicity_2002; @hwang_body_2017; @ostling_visual_2018; @thompson_defining_2019; @voronin_osnovy_2006; @flaksman2020stanislav). This is more common and straightforward for sign languages than for spoken languages [@thompson_defining_2019]. It is usually done manually, but Östling et al. -@ostling_visual_2018 have shown for sign languages that (for some mappings at least) this can also be automated. Such studies usually involve cross-linguistic data, and aim to uncover universalities in how different languages map sound to meaning--that is, *where* iconicity comes from.

A wider variety of approaches target iconicity *indirectly* through related phenomena. For example, data-driven approaches (e.g. @wichmann_sound_2010; @blasi_soundmeaning_2016; @joo_phonosemantic_2019; @johansson_typology_2020; @winter_trilledr) use comparisons across unrelated languages to identify form-meaning correspondences that occur at rates higher than chance. This requires large amounts of parallel data from distinct phyla, and existing studies usually focus on basic vocabulary--where such data is most readily available. Like descriptive approaches, data-driven approaches identify *specific* form-meaning associations involved in iconic effects, but with the added advantage of not needing to rely on the intuitions and manual labour of individual researchers. However, they *do* rely on the existence of translation equivalents in multiple languages for the words under study. This can pose a particular challenge for highly iconic words, as these often also have highly specific meanings (@lupyan_language_2018; @akita_toward_2012). Thus, these approaches are less likely to be helpful for those with *specific* words for which iconicity measures are required.

Another body of work which capitalises on big data involves the study of *affective iconicity* [e.g. @ullrich_phonological_2016; @aryani_why_2018]. These studies use large datasets of affective norms to identify phonological segments loaded with affective meaning (by taking average affectiveness values across all words containing those segments, and identifying segments with values deviating from the global neutral mean). Although limited to associations involving affect, this approach has been successfully used to investigate iconic effects in language processing [@schmidtke_effects_2018; Ullrich in prep; @aryani_affective_2018], and has even been used to explain the affective meaning of poems [@ullrich_relation_2017].

@@

In iconicity rating tasks, the concept of iconicity is explained to participants, and participants are asked to rate the iconicity of given form-meaning pairs directly [e.g. @vinson_british_2008; @perry_iconicity_2015; @winter_which_2017; @occhino_iconicity_2017; @thompson_iconicity_2020; @punselie_iconicity_2020]. Measurements from these studies have been found to correlate in meaningful ways with factors related to iconicity--e.g. encoding of sensory information [@winter_which_2017], semantic domains [@punselie_iconicity_2020], semantic neighbourhood density [@sidhu_lonely_2018], word class [@perry_iconicity_2015; @thompson_iconicity_2020], age of acquisition [@vinson_british_2008; @perry_iconicity_2015], facilitated processing [@thompson_link_2009; @occhino_role_2020], funniness and structural markedness [@dingemanse_playful_2020]. However, ratings given are influenced by the linguistic experience of participants (@occhino_iconicity_2017; @sevcikova_sehyr_perceived_2019), so some care is needed to ensure measures are taken from appropriate participant groups. Further, ratings from native speakers may also capture non-iconic regularities like semantic transparency and indexicality [@dingemanse_playful_2020; @winter_iconicity_2021], so cannot be taken as direct reflections of degree of iconicity.

In guessing experiments, participants are asked to guess the meaning of foreign/novel signs from a choice of two or more alternatives [e.g. @kohler_gestalt_1929; @tsuru_problem_1933; @brown_phonetic_1955; @kunihira_effects_1971; @ramachandran_synaesthesia_2001; @dingemanse_what_2016; @tzeng_specificity_2017; @perlman_iconicity_2015; @perlman_people_2018; @lockwood_sound-symbolism_2016]. Signs whose meanings are correctly guessed at rates higher than chance are considered iconic. The guessability of signs has also been linked to facilitated processing [@ormel_role_2009] and learnability [@perlman_people_2018; @lockwood_sound-symbolism_2016], and correlated with iconicity ratings [@punselie_iconicity_2020]. However, some signs that were given low iconicity ratings were actually highly guessable, while there were also a few signs that, despite receiving high ratings, were not easily guessed  [@punselie_iconicity_2020: 23]. This suggests that guessing and rating tasks may tap into slightly different constructs, or be sensitive to experimental design constraints.

@@

The data from the study, as well as the code used to produce the experiments and analyses can be found on the Open Science Framework, at https://osf.io/j57uc/?view_only=d59ffde4bc42467fafb71a904bef8d14.

@@

The stimuli were 304 words taken from the *Japonic Sensory Lexicon* database [@McLean_JaponicSensoryLex_2022], which is a collection of words for 110 sensory concepts in 49 Japonic varities. The words in the database were collated from three different sources: Sanada and Tomosada [-@sanada_kenbetsu_2015; -@sanada_kenbetsu_2018], and Hirayama [-@hirayama_gendai_1992]. Any available concepts related to the senses were included. For all three sources, words for each concept were collected from speakers of different Japonic varieties by asking the speakers to translate sentences from the standard variety into their local variety. Further description of the concepts and the sentences used to elicit them can be found in the Supplementary Materials

@@

graph%>%
  mutate(domain=fct_relevel(domain,c("sound","action","shape & size","appearance","texture","temperature","taste","pain","bodily feeling","emotion")))%>%
  ggplot(aes(x=domain,y=n,fill=stratum))+geom_col()+theme_minimal()+labs(x="Semantic domain",fill="Lexical stratum")+theme(axis.text.x = element_text(angle=15))+scale_fill_manual(values=c("darkgrey","black"))+ coord_fixed(ratio = 0.1)->plot
```

@@

The number of these words which were guessed significantly above chance (with a sample of 30 guesses per word, an accuracy greater than 2/3 is needed to be sure the true accuracy is above 0.5) was higher in the experiment with opposite foils than in the experiment with random foils. Using opposite foils, half of the words in the sample (15/30) were guessed significantly above chance, whereas with the random foils this number fell to just 1 in 5 words (6/30). Only two words were guessed significantly above chance in the random foil condition, but not in the opposite foil condition. However, both were guessed numerically above chance in the opposite foil condition as well (with accuracies of 52% and 57%). The differences between the two conditions are shown in Figure \@ref(fig:sensitivitycheck). A chi-squared test confirmed the differences to be significant (χ2(1)=4.69,p=.03).

@@

For both ideophones and prosaic words, there was a strong correlation between the ratings and the guesses. This correspondence was slightly better for ideophones (r = 0.66, 95% CI [0.44, 0.79], 95% CI [0.53, 0.76], t(99) = 8.68, p < .001) compared to prosaic words (r = 0.41, 95% CI [0.29, 0.52], t(201) = 6.45, p < .001). To explore this relationship further, we created two linear regression models: one predicting guesses from ratings, lexical stratum, and the interaction between these two factors, and a second model predicting ratings from guesses, lexical stratum, and the interaction between these two factors. Both models again used z-scores rather than raw ratings and guesses. The models were created using the R package `lmerTest` (version `r packageVersion("lmerTest")`) [@lmertest]. The output of the first model (predicting guesses from ratings) is shown in Table \@ref(tab:guessesmodel).

@@

For both ideophones and prosaic words, we found that the guesses and ratings from non-speakers were strongly correlated in the *same way*. This suggests that the two measures are indeed tapping into *something* in common, which we suggest are form-meaning associations. It's notable that these associations are found acr

in both ideophones *and* prosaic words (even if they are more common in the former), and that there are also ideophones that participants had no strong associations with--highlighting that there is no straightforward relationship between ideophones and iconicity; variable levels of iconicity can be found in both ideophones and prosaic words [see also @dingemanse_what_2016; @brown_phonetic_1955; @kunihira_effects_1971].

However, ideophones were consistently rated *higher* in iconicity than non-ideophones—even when guessed at the same accuracies. We believe this is due to the majority of ideophones in the dataset being reduplicated. The two lowest rated ideophones–*sappari* and *syippori*–were among the few in the data that do not use a reduplicated template. Previous work has found a relationship between iconiticity ratings and structural markedness (Dingemanse and Thompson 2020). Ordinarily when we use words, the form of the words is not important--only their content. However, when iconicity is involved (and also funniness, as this paper found) the form of the words *is* important, and structural markedness may serve as a metacommunicative cue to signal this. The results from the rating task in this study suggest that structural markedness (in this case, reduplication) does indeed lead to the perception of iconicity. Interestingly, however, it does not seem to confer any advantage--over and above that provided by associations between form and meaning--in the guessing task. What the ratings seem to be picking up on is both form-meaning associations *and* structural markedness. If they were only picking up on structural markedness, then we would not expect the ratings to vary with the guesses in the same way for both ideophones and non-ideophones. However, if they were only picking up on form-meaning associations (like the guesses seem to), then we would not expect ideophones to consistently receive higher iconicity ratings, even when guessed at the same accuracies.

While ideophones and non-ideophones with high iconicity ratings were generally guessed equally well, there was an interaction effect such that ideophones with low iconicity ratings were guessed *worse* than non-ideophones with comparable ratings. Low iconicity ratings in non-ideophones generally correspond to chance guessing accuracy--reflecting a lack of associations--whereas equivalent ratings in ideophones correspond to *below chance* guessing accuracy--reflecting *negative associations* masked in the rating results by the tendency to inflate ratings for ideophones. The result highlights the subjective nature of form-meaning associations [@occhino_iconicity_2017], and suggest that by taking above-chance guessing accuracy as a measure of iconicity we are only picking up on those associations that happen to correspond--and crucially not to conflict--between (in this case) English and Japanese speakers. This is something researchers should be aware of, because it means that decisions about *whom* to collect iconicity measurements from are very important.

Poor guessing accuracies could be explained by conflicting associations. For example, one person commented that *zarazara* as a word for ‘ROUGH’ sounds wrong to them because they associate the word *zara* with the clothing store ZARA, and clothes are soft. These could be masking either a lack of resemblance between form and meaning, or a resemblance between form and meaning that is ignored in the face of the conflicting association. Of course, conversely one could argue that high guessing accuracies could then *also* be explained by non-resemblance based associations. Fortunately, the chances of English speakers having the *same* form-meaning association as Japanese speakers based on something random (like the name of a clothing store) should be much smaller than the chances of them having a *different* association. Thus, if they do exist, we predict that the majority of non-resemblance based associations should be reflected in poorer rather than higher guessing accuracies. It is interesting that conflicting associations appear to have a stronger effect in ideophones (some of which were guessed at accuracies significantly below chance) than in non-ideophones (which were guessed below chance only rarely). Perhaps because participants are more likely to view form-meaning relationships in ideophones as meaningful (due to their structural markedness), they are more likely to attach conflicting associations to ideophones, leading to a stronger sense that the ideophone should correspond to the *opposite* of the given meaning.

Finally, the comparison between guesses and ratings highlights a bias in the design of the rating task, as all the words in the data were in the end rated as "slightly" iconic (even by non-speakers), whereas the guessing results show that a good number of these were not guessed any better than chance. This is probably an artefact of the design of the rating task, in which "arbitrary" is only 1 of the 7 or 10 points on the scale--the rest indicating varying degrees of form-meaning resemblance. This is not so problematic, as the ratings still correlate with guessing accuracy, indicating that the relative differences in iconicity still hold. However, it highlights that interpretations of particularly *low* iconicity ratings as reflections of iconicity should be viewed with caution.

To sum up, by directly comparing iconicity ratings and guessing accuracies for both ideophones and non-ideophones, we are able to better understand how these measures relate to each other and to iconicity. Differences between iconicity ratings of ideophones and non-ideophones--even when guessability is the same--highlight the role of structural markedness in enhancing perceived iconicity. Consistencies between guesses and ratings within these word types reassure us that, despite this, form-meaning association biases play a large role in explaining each measure. The unexpected result of ideophones also being poorly guessed reminds us that these biases are just that--subjective--and thus, while good guessing accuracies can be interpreted as positive evidence of language-external iconicity, average or poor guessing accuracies *cannot* be interpreted as negative evidence of iconicity of either kind (language-internal or external). Similarly, while high ratings are likely to be a good indication of iconicity, researchers should be cautious of taking lower ratings indicating "slight" iconicity at face value.

@@

Given the limitations of each measure discussed above, for a full picture of iconicity in the lexicon, a synthesis of measures is recommended. Iconicity ratings from native speakers can tell us which mappings are meaningful to them, while comparisons with guessability for non-speakers can establish which of these mappings are likely to be meaningful on their own, versus which *become* meaningful through experience with the rest of the lexicon (language-external versus language-internal iconicity). @occhino_role_2020 found differential effects of these two types of mappings in sign processing by signers of different proficiencies. They found that both types of iconicity were helpful to signers of a lower proficiency, but that for highly proficient signers only the most pervasive, language-external mappings made a difference to processing.

Since structural markedness appears to boost ratings but not guessability, the two measures could also be used together to tease apart the influence of structural markedness versus form-meaning resemblances in driving iconic effects. Iconic effects are generally assumed to be driven by form-meaning resemblances, but since these often go together with structural markedness it would be useful to know the relative contribution of each factor in different situations and tasks. Since structural markedness is also subjective and language-relative, the choice of participants used in the rating task will be important. Finally, having guesses alongside ratings would allow researchers to make better decisions about the cut-off point between "high" and "low" iconicity ratings, as these could be motivated by correlations with guessing accuracy, rather than e.g. arbitrarily dividing the scale at the half-way point, or using percentiles tied to particular datasets.

Guesses and ratings could also be used to inform results from descriptive or data-driven measures, and vice-versa. For example, the psychological reality and relative *strength* of form-meaning mappings described in lexicons could be tested using guessing or rating tasks with nonwords [e.g. @kwon_empirically_2017]. Similarly, descriptive (semiotic) analysis, along with cross-linguistic comparisons where possible could be used to understand and explain why some mappings are stronger or more universal than others.

## Methodological improvements and evaluation

We also sought to evaluate different rating and guessing paradigms, and were able to make some methodological improvements--particularly to the guessing design. We contrasted two different guessing paradigms--one where participants are given a word and have to match it to the correct meaning from two translations, and another where participants are given a meaning and have to match it to the correct word from two words. Content-wise, choosing between words is more comparable to the design of a rating task, as both give the meaning first and involve only one English translation. This is important as the use of translations leads to several complications in the experiment design. First, English-speaking participants can be expected to have more associations with English words than with unknown Japanese words, which could make the task more sensitive to the particular translations chosen. In addition, care has to be taken when choosing foils as well, as they should preferably be from the same semantic domain, of a similar length of characters, and not synonyms or antonyms of the correct translation. This makes the guessing between translations design more difficult to use, particularly in small semantic domains (e.g. taste).

In contrast, a task in which participants guess between words is much easier to operationalise. This design can be easily adopted to any semantic domain or any kind of data, and the foils can be chosen either randomly from the other trials or selected based on specific criteria. In our case we chose foils that were phonologically as distinct as possible from the word being tested. The first reason for this was theoretical: in a task designed to pick up on form-meaning resemblance, words that sound very different should not be suited to that meaning, enhancing the effect of iconicity. The second was practical: our stimuli had artificially flat intonations, which has been shown to decrease performance in guessing studies [@kunihira_effects_1971; @dingemanse_what_2016]. By creating opposite sounding words, we tried to increase performance in the guessing task despite the degraded stimuli.

When comparing results from the two designs, choosing between words is again the more favourable of the two. First, as expected changing the translation or foil used led to less inconsistent results when choosing between words compared to when choosing between translations. Second, the range of guessing scores is much wider when choosing between words compared to choosing between translations, with more words guessed at high accuracy levels, and fewer words around chance level. This is especially true when the foils used are maximally phonologically distinct. However, for datasets that are relatively higher in iconicity to begin with--e.g. gestures or hand signs--using random foils will likely already produce good results. The better results in the choosing between words design can be easily explained, as the task was comparatively easier than in the choosing between translations design, where participants were restricted to choosing between two similar options (non-antonyms from the same semantic domain). The options in the choosing between words design are more distinct--even maximally so. In sum, the choosing between words design is both more robust, more sensitive to iconicity, and more discriminating between lesser and greater amounts of iconicity, so this is the design that we recommend and that we have implemented in `icotools`.

>>>>

---
title: "Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon"
output:
  bookdown::html_document2:
 #   reference_docx: wordstyles.docx
    toc: true
    toc_float: true
    number_sections: true
    css: custom.css
    code_folding: hide
always_allow_html: true
bibliography: sources.bib
csl: apa.csl

@@

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(report)
library(lmerTest)
library(plotly)
library(corrr)
library(boot)
library(zeallot)
library(pander)
library(sjPlot)
library(ggeffects)
library(gridExtra)
library(gghalves)
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)
```

@@

Iconicity--defined here as a perceived resemblance between aspects of form and meaning--is now increasingly recognised as a *key property of language* [see e.g. @dingemanse_arbitrariness_2015; @perniss_iconicity_2010; @perniss_bridge_2014 for a review], raising questions about where iconic mappings come from, and what they do. To answer these questions, we first need to be able to *identify* iconicity in language and, ideally, to *measure* it.

@@

With descriptive approaches, iconicity as a theoretical concept is coded for directly [e.g. @pietrandrea_iconicity_2002; @hwang_body_2017; @ostling_visual_2018; @thompson_defining_2019; @voronin_osnovy_2006; @flaksman2020stanislav]. This is more common and straightforward for sign languages than for spoken languages [@thompson_defining_2019]. It is usually done manually, but Östling et al. [-@ostling_visual_2018] have shown for sign languages that (for some mappings at least) this can also be automated. Such studies usually involve cross-linguistic data, and aim to uncover universalities in how different languages map sound to meaning--that is, *where* iconicity comes from.

A wider variety of approaches target iconicity *indirectly* through related phenomena. For example, data-driven approaches [e.g. @wichmann_sound_2010; @blasi_soundmeaning_2016; @joo_phonosemantic_2020; @johansson_typology_2020; @winter_trilledr] use comparisons across unrelated languages to identify form-meaning correspondences that occur at rates higher than chance. This requires large amounts of parallel data from distinct phyla, and existing studies usually focus on basic vocabulary--where such data is most readily available. Like descriptive approaches, data-driven approaches identify *specific* form-meaning associations involved in iconic effects, but with the added advantage of not needing to rely on the intuitions and manual labour of individual researchers. However, they *do* rely on the existence of translation equivalents in multiple languages for the words under study. This can pose a particular challenge for highly iconic words, as these often also have highly specific meanings [@lupyan_language_2018; @akita_toward_2012]. Thus, these approaches are less likely to be helpful for those with *specific* words for which iconicity measures are required.

Another body of work which capitalises on big data involves the study of *affective iconicity* [e.g. @ullrich_phonological_2016; @aryani_why_2018]. These studies use large datasets of affective norms to identify phonological segments loaded with affective meaning (by taking average affectiveness values across all words containing those segments, and identifying segments with values deviating from the global neutral mean). Although limited to associations involving affect, this approach has been successfully used to investigate iconic effects in language processing [@schmidtke_effects_2018; @aryani_affective_2018], and has even been used to explain the affective meaning of poems [@ullrich_relation_2017].

@@

In iconicity rating tasks, the concept of iconicity is explained to participants, and participants are asked to rate the iconicity of given form-meaning pairs directly [e.g. @vinson_british_2008; @perry_iconicity_2015; @winter_which_2017; @occhino_iconicity_2017; @thompson_iconicity_2020; @punselie_iconicity_2020]. Measurements from these studies have been found to correlate in meaningful ways with factors related to iconicity--e.g. encoding of sensory information [@winter_which_2017], semantic domains [@punselie_iconicity_2020], semantic neighbourhood density [@sidhu_lonely_2018], word class [@perry_iconicity_2015; @thompson_iconicity_2020], age of acquisition [@vinson_british_2008; @perry_iconicity_2015], facilitated processing [@thompson_link_2009; @occhino_role_2020], funniness and structural markedness [@dingemanse_playful_2020]. However, ratings given are influenced by the linguistic experience of participants [@occhino_iconicity_2017; @sevcikova_sehyr_perceived_2019], so some care is needed to ensure measures are taken from appropriate participant groups. Further, ratings from native speakers may also capture non-iconic regularities like semantic transparency and indexicality [@dingemanse_playful_2020; @winter_iconicity_2021], so cannot be taken as direct reflections of degree of iconicity.

In guessing experiments, participants are asked to guess the meaning of foreign/novel signs from a choice of two or more alternatives [e.g. @kohler_gestalt_1929; @tsuru_problem_1933; @brown_phonetic_1955; @kunihira_effects_1971; @ramachandran_synaesthesia_2001; @dingemanse_what_2016; @tzeng_specificity_2017; @perlman_iconicity_2015; @perlman_people_2018; @lockwood_sound2016]. Signs whose meanings are correctly guessed at rates higher than chance are considered iconic. The guessability of signs has also been linked to facilitated processing [@ormel_role_2009] and learnability [@perlman_people_2018; @lockwood_sound2016], and correlated with iconicity ratings [@punselie_iconicity_2020]. However, some signs that were given low iconicity ratings were actually highly guessable, while there were also a few signs that, despite receiving high ratings, were not easily guessed  [@punselie_iconicity_2020: 23]. This suggests that guessing and rating tasks may tap into slightly different constructs, or be sensitive to experimental design constraints.

@@

The data from the study, as well as the code used to produce the experiments and analyses can be found on the Open Science Framework, at https://osf.io/j57uc/?view_only=0f15b81806434bc99c29566d68f3f715.

@@

The stimuli were 304 words taken from the *Japonic Sensory Lexicon* database [AUTHOR 2022], which is a collection of words for 110 sensory concepts in 49 Japonic varities. The words in the database were collated from three different sources: Sanada and Tomosada [-@sanada_kenbetsu_2015; -@sanada_kenbetsu_2018], and Hirayama [-@hirayama_gendai_1992]. Any available concepts related to the senses were included. For all three sources, words for each concept were collected from speakers of different Japonic varieties by asking the speakers to translate sentences from the standard variety into their local variety. Further description of the concepts and the sentences used to elicit them can be found in the Supplementary Materials.

@@

graph%>%
  mutate(domain=fct_relevel(domain,c("sound","action","shape & size","appearance","texture","temperature","taste","pain","bodily feeling","emotion")))%>%
  ggplot(aes(x=domain,y=n,fill=stratum))+geom_col()+theme_minimal()+labs(x="Semantic domain",fill="Lexical stratum")+theme(axis.text.x = element_text(angle=15))+scale_fill_manual(values=c("darkgrey","black"))+ coord_fixed(ratio = 0.1)
```

@@

The number of these words which were guessed significantly above chance (with a sample of 30 guesses per word, an accuracy greater than 2/3 is needed to be sure the true accuracy is above 0.5) was higher in the experiment with opposite foils than in the experiment with random foils. Using opposite foils, half of the words in the sample (15/30) were guessed significantly above chance, whereas with the random foils this number fell to just 1 in 5 words (6/30). Only two words were guessed significantly above chance in the random foil condition, but not in the opposite foil condition. However, both were guessed numerically above chance in the opposite foil condition as well (with accuracies of 52% and 57%). The differences between the two conditions are shown in Figure \@ref(fig:sensitivitycheck). A chi-squared test confirmed the differences to be significant (*χ2(1)=4.69,p=.03*).

@@

For both ideophones and prosaic words, there was a strong correlation between the ratings and the guesses. This correspondence was slightly better for ideophones (*r = 0.66, 95% CI [0.44, 0.79], 95% CI [0.53, 0.76], t(99) = 8.68, p < .001*) compared to prosaic words (*r = 0.41, 95% CI [0.29, 0.52], t(201) = 6.45, p < .001*). To explore this relationship further, we created two linear regression models: one predicting guesses from ratings, lexical stratum, and the interaction between these two factors, and a second model predicting ratings from guesses, lexical stratum, and the interaction between these two factors. Both models again used z-scores rather than raw ratings and guesses. The models were created using the R package `lmerTest` (version `r packageVersion("lmerTest")`) [@lmertest]. The output of the first model (predicting guesses from ratings) is shown in Table \@ref(tab:guessesmodel).

@@

For both ideophones and prosaic words, we found that the guesses and ratings from non-speakers were strongly correlated in the *same way*. This suggests that the two measures are indeed tapping into *something* in common, which we suggest are form-meaning associations. It’s notable that these associations are found in varying amounts across both strata of the lexicon, highlighting that there is no straightforward relationship between ideophones and iconicity; variable levels of iconicity can be found in both ideophones and prosaic words [see also @dingemanse_what_2016; @kunihira_effects_1971].

However, ideophones were consistently rated *higher* in iconicity than prosaic words—even when guessed at the same accuracies. We believe this is due to the majority of ideophones in the dataset being reduplicated. The two lowest rated ideophones–*sappari* ‘clean’ and *guttari ‘*exhausted*’*–were among the few in the data that do not use a reduplicated template. Previous work has found a relationship between iconicity ratings and structural markedness [@dingemanse_playful_2020]. Ordinarily when we use words, the form of the words is not important–only their content. However, when iconicity is involved the form of the words *is* important, and structural markedness may serve as a metacommunicative cue to signal this. The results from the rating task suggest that structural markedness (in this case, reduplication) leads to the perception of iconicity. Interestingly, however, it does not seem to confer an advantage–over and above that provided by associations between form and meaning–in the guessing task. What the ratings seem to be picking up on is both form-meaning associations *and* structural markedness. If they were only picking up on structural markedness, then we would not expect the ratings to vary with the guesses in the same way for both ideophones and prosaic words. However, if they were only picking up on form-meaning associations (like the guesses seem to), then we would not expect ideophones to consistently receive higher iconicity ratings, even when guessed at the same accuracies.

While ideophones and prosaic words with high iconicity ratings were generally guessed equally well, there was an interaction effect such that ideophones with low iconicity ratings were guessed *worse* than prosaic words with comparable ratings. Low iconicity ratings in prosaic words generally correspond to chance guessing accuracy–reflecting a lack of associations–whereas equivalent ratings in ideophones correspond to *below chance* guessing accuracy–reflecting *negative associations* masked in the rating results by the tendency to inflate ratings for ideophones. The result highlights the subjective nature of form-meaning associations [@occhino_iconicity_2017], and suggests that by taking above-chance guessing accuracy as a measure of iconicity we are only picking up on those associations that happen to correspond–and crucially not to conflict–between (in this case) English and Japanese speakers. This is something researchers should be aware of, because it means that decisions about *whom* to collect iconicity measurements from are very important.

Poor guessing accuracies could be explained by conflicting associations. For example, one person commented that *zarazara* as a word for ‘ROUGH’ sounds wrong to them because they associate the word *zara* with the clothing store ZARA, and clothes are soft. These could be masking either a lack of resemblance between form and meaning, or a resemblance between form and meaning that is missed in the face of the conflicting association. Of course, conversely one could argue that high guessing accuracies could then *also* be explained by non-resemblance based associations. Fortunately, the chances of English speakers having the *same* form-meaning association as Japanese speakers based on something random (like the name of a clothing store) should be much smaller than the chances of them having a *different* association. Thus, if they do exist, we predict that the majority of non-resemblance based associations should be reflected in poorer rather than higher guessing accuracies.

It is interesting that conflicting associations appear to have a stronger effect in ideophones (some of which were guessed at accuracies significantly below chance) than in prosaic words (which were guessed below chance only rarely). Perhaps because participants are more likely to view form-meaning relationships in ideophones as meaningful (due to their structural markedness), they are more likely to attach conflicting associations to ideophones, leading to a stronger sense that the ideophone should correspond to the *opposite* of the given meaning [for more on this see discussion in @webster_2017; @akita_ideophone_2019: 19]. This is supported by the larger person-total correlations reported for ideophones compared to prosaic words in Section 3.3; that participants were more consistent with each other when rating ideophones suggests that intuitions about these words were more pervasive.

Finally, the comparison between guesses and ratings highlights a bias in the design of the rating task. All the words in the data were rated  “slightly” iconic, even though the guessing results show a good number of these were not guessed any better than chance. This is probably an artefact of the scale used, in which arbitrariness is represented by only 1 of the 7 points, while the other 6 all indicate varying degrees of form-meaning resemblance. This is not so problematic, as the ratings still correlate with guessing accuracy, indicating that the relative differences in iconicity still hold. However, it highlights that interpretations of particularly *low* iconicity ratings as reflections of iconicity should be viewed with caution.

To sum up, by directly comparing iconicity ratings and guessing accuracies for both ideophones and prosaic words, we are able to better understand how these measures relate to each other and to iconicity. Differences between iconicity ratings for ideophones and prosaic words–even when guessability is the same–highlight the role of structural markedness in enhancing perceived iconicity. Consistencies between guesses and ratings within these word types reassure us that, despite this, form-meaning association biases play a large role in explaining each measure. The unexpected result of ideophones also being poorly guessed reminds us that these biases are just that–subjective–and thus, while good guessing accuracies can be interpreted as positive evidence of language-external iconicity, average or poor guessing accuracies *cannot* be interpreted as negative evidence of iconicity of either kind (language-internal or external). Similarly, while high ratings are likely to be a good indication of iconicity, researchers should be cautious of taking lower ratings indicating “slight” iconicity at face value.

@@

Given the limitations of each measure discussed above, for a full picture of iconicity in the lexicon, a synthesis of measures is recommended. Iconicity ratings from native speakers can tell us which mappings are meaningful to them, while comparisons with guessability for non-speakers can establish which of these mappings are likely to be meaningful on their own, versus which *become* meaningful through experience with the rest of the lexicon. Occhino et al. [-@occhino_role_2020] refer to this distinction as language-external versus language-internal iconicity. It is a distinction with real correlates, as their study found differential effects for each type of iconicity on sign language processing.

Since structural markedness appears to boost ratings but not guessability, the two measures could be used together to tease apart the influence of structural markedness versus form-meaning resemblances in driving iconic effects. Iconic effects are generally assumed to be driven by form-meaning resemblances, but as these often go together with structural markedness it would be useful to know the relative contribution of each factor in different situations and tasks. Having guesses alongside ratings would also allow researchers to make better decisions about the cut-off point between “high” and “low” iconicity ratings, as these could be motivated by correlations with guessing accuracy, rather than e.g. arbitrarily dividing the scale at the half-way point, or using percentiles tied to particular datasets.

Guesses and ratings could also be used to inform results from descriptive or data-driven measures, and vice-versa. For example, the psychological reality and relative *strength* of form-meaning mappings described in lexicons could be tested using guessing or rating tasks with nonwords [e.g. @kwon_empirically_2017]. Similarly, descriptive (semiotic) analysis, along with cross-linguistic comparisons where possible, could be used to understand where these associations come from and explain why some are stronger or more universal than others.

## Methodological contributions

As a final contribution of this study, having evaluated different guessing and rating paradigms, we were able to make some methodological improvements—particularly to the guessing design. We contrasted two different guessing paradigms–one where participants are given a word and asked to match it to the correct meaning from two translations, and another where participants are given a meaning and have to match it to the correct word from two words. Content-wise, choosing between words is more comparable to the design of the rating task, as both give the meaning first and involve only a single English translation. This is important as the use of translations leads to several complications in the interpretation of guessing results. First, English-speaking participants can be expected to have more associations with English words than with unknown Japanese words, which could make the task more sensitive to the particular translations chosen. In addition, care has to be taken when choosing foils, as they should preferably be from the same semantic domain, of a similar length of characters, and neither synonyms nor antonyms of the correct translation (see description in Section 2.3). This makes the guessing between translations design more difficult to implement, particularly for small semantic domains (e.g. taste).

In contrast, a task in which participants guess between words is much easier to operationalise. This design can be easily adopted to any semantic domain or type of data, with foils either chosen randomly from among other items in the task, or specified using motivated criteria. In our case, we used foils that were maximally phonologically distinct from the word being tested. Our logic was that if the real word is a good fit for the meaning, then a word that sounds very different to it should be a bad fit for the meaning. So using opposite-sounding foils enhances the task’s sensitivity to iconicity.

When comparing results between the two designs, the meaning-to-word design clearly emerges as the more favourable of the two. First, as expected changing the translation or foil used led to less inconsistent results when choosing between words compared to when choosing between translations (Section 3.1). Second, the range of guessing scores is much wider in the meaning-to-word design compared to the word-to-meaning design, with more words guessed at high (and low) accuracy levels, and fewer words guessed at chance (Section 3.2). This is especially true when the foils used are maximally phonologically distinct. The same effect can be seen in form-to-meaning designs when comparing baseline differences in guessing accuracies between studies that used antonyms versus random words as foils [e.g. @lockwood_sound2016 versus @dingemanse_what_2016]. In sum, our new meaning-to-word guessing design is both more robust to different choices of translations and foils, more sensitive to iconicity, and more discriminating between lesser and greater amounts of iconicity.

We also considered the effects of using non-native speakers versus native speakers as raters in the rating task. We found that using native speakers provided a greater spread of measures (Section 3.2), probably because native speakers’ familiarity with the words makes them more confident in giving high or low ratings. However, we did not find any difference in the consistency of the ratings provided between the two groups—non-native speakers were just as consistent as native speakers in rating the iconicity of Japanese words. Since both types of ratings are fairly reliable, we recommend choosing raters according to what makes the most theoretical sense in regards to the question being asked. For example, to predict language processing in adult native speakers, it makes sense to also use native speakers as raters; to predict outcomes in second language acquisition, using non-speakers is more theoretically sound. The important thing to remember is that perceptions of iconicity are subjective and experience dependent, and this needs to be accounted for when collecting iconicity measures [@occhino_iconicity_2017; @occhino_role_2020].

We also created a python package, icotools, which can be used to automatically generate guessing and rating tasks from wordlists, following the designs described in Section 2.4 and Section 2.5. A full discussion of icotools and its functionality is provided in the Supplementary Materials.  Some of its main advantages are (1) the ability to generate both guessing and rating tasks from the same wordlist, (2) the ability to specify multiple translations and multiple foils per word, so that results cannot be unduly influenced by one particular choice of translation or foil, and (3) the ability to include a set of ‘control items’ in every experiment, as an additional way to monitor whether participants are behaving as expecting (see Section 2.2). As well as audio files, the package supports videos and images as stimuli formats, making it suitable for investigations of iconicity in multiple modalities.
